---
title: "Gaussian Process regression - R laGP (WIP)"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<br>

A replication of the Gaussian Process regression implementation lifted from chapter 5 of [Surrogates](https://bookdown.org/rbg/surrogates/chap5.html), and an application of the same code to real data.

<br>

## Chapter example

<br>

### Data

Create some dummy data being an independent and dependent variable, along with a grid of the independent variable values.

```{r }
# Training data
n <- 8
X <- matrix(seq(0, 2*pi, length=n), ncol=1)         # independent variable
y <- sin(X) 

# Predictive grid
XX <- matrix(seq(-0.5, 2*pi + 0.5, length=100), ncol=1)

# Nugget / jitter
eps <- sqrt(.Machine$double.eps)             

```

<br>

### Using function `newGP`

Investigate parameters:

d: a positive scalar lengthscale parameter for an isotropic Gaussian correlation function (newGP); or a vector for a separable version (newGPsep), and

g: a positive scalar nugget / jitter parameter

```{r }
library('laGP')

# Derive sensible parameter ranges
d <- darg(NULL, X)

y[1] <- 1e-3 # negate garg error: Error in check.arg(g) : d$min should be a postive scalar < d$max
g <- garg(NULL, y)

par(mfrow=c(1,2))

for(parm in c('d','g')) {
  # GP
  gpi <- newGP(X, y, d=d$start, g=g$start*0.5, dK=TRUE)
  #mle <- mleGP(gpi) 
  mle <- mleGP(gpi, param=parm, tmin=ifelse(parm == 'd', d$min, g$min), tmax=ifelse(parm == 'd', d$max, g$max))
  p <- predGP(gpi, XX)
  YY <- mvtnorm::rmvnorm(100, p$mean, p$Sigma)
  q1 <- p$mean + qnorm(0.05, 0, sqrt(diag(p$Sigma)))
  q2 <- p$mean + qnorm(0.95, 0, sqrt(diag(p$Sigma)))
  
  # Plot
  matplot(XX, t(YY), type="l", col="gray", lty=1, 
         main = paste0("Optimising ", ifelse(parm == 'd', 'lengthscale', 'nugget / jitter')), 
         xlab="x", ylab="sin(x)", bty = "n")
  points(X, y, pch=20, cex=2)
  lines(XX, p$mean, lwd=2)
  lines(XX, sin(XX), col="blue")
  lines(XX, q1, lwd=2, lty=2, col=2)
  lines(XX, q2, lwd=2, lty=2, col=2)
  
  deleteGP(gpi)
  rm(mle)
}
```

<br>

### Using function `newGPsep`

Separable GP, the lengthscale parameter is allowed to decay at differing rates dependent on the input data. It is anisotropic, rather than isotropic.

```{r }
gpi <- newGPsep(X, y, d=d$start, g=g$start*0.5, dK=TRUE)
mle <- mleGPsep(gpi, param='both', tmin=c(d$min, g$min), tmax=c(d$max, g$max))
p <- predGPsep(gpi, XX)
YY <- mvtnorm::rmvnorm(100, p$mean, p$Sigma)
q1 <- p$mean + qnorm(0.05, 0, sqrt(diag(p$Sigma)))
q2 <- p$mean + qnorm(0.95, 0, sqrt(diag(p$Sigma)))

# Plot
matplot(XX, t(YY), type="l", col="gray", lty=1, 
       main = "seperable GP ", 
       xlab="x", ylab="sin(x)", bty = "n")
points(X, y, pch=20, cex=2)
lines(XX, p$mean, lwd=2)
lines(XX, sin(XX), col="blue")
lines(XX, q1, lwd=2, lty=2, col=2)
lines(XX, q2, lwd=2, lty=2, col=2)

deleteGPsep(gpi)
rm(mle)
```

<br>

## Stock data example

<br>

The procedures above are now applied to real data, that being stock valuations and fundamentals.

<br>

### Data

```{r, results="hide", message=FALSE, warning=FALSE}
library(romerb)
data("stock_data")
fundamental_raw <- stock_data
rm(stock_data)

# Medical devices sector only
df <- fundamental_raw[fundamental_raw$sector == 7, ]   
df$log_mkt_cap <- log(df$mkt_cap)
df$log_book <- log(-df$total_equity_cln)
df$roe <- -df$roe
df <- df[df$date_stamp == as.Date('2021-06-30'), c('log_book','log_mkt_cap','log_pb','roe','leverage')]

# nugget / jitter
eps <- sqrt(.Machine$double.eps)

# Training data
X <- matrix(df$roe)
y <- matrix(df$log_pb)

# Predictive grid
XX <- matrix(seq(min(X), max(X), length=200), ncol=1)

# Plot
plot(X, y, main = "Medical devices sector",
     xlab = "Return on equity", ylab = "Price book ratio",
     pch = 19, frame = FALSE)
```

```{r, results="hide", message=FALSE, warning=FALSE}
d <- darg(NULL, X)
g <- garg(NULL, y)

gpi <- newGPsep(X, y, d=d$start, g=g$start*0.5, dK=TRUE)
mle <- mleGPsep(gpi, param='both', tmin=c(d$min, g$min), tmax=c(d$max, g$max))
p <- predGPsep(gpi, XX)
YY <- mvtnorm::rmvnorm(100, p$mean, p$Sigma)
q1 <- p$mean + qnorm(0.05, 0, sqrt(diag(p$Sigma)))
q2 <- p$mean + qnorm(0.95, 0, sqrt(diag(p$Sigma)))

# Plot
matplot(XX, t(YY), type="l", col="gray", lty=1, 
       xlab = "Return on equity", ylab = "Price book ratio", bty = "n")
points(X, y, pch=20, cex=2)
lines(XX, p$mean, lwd=2)
lines(XX, q1, lwd=2, lty=2, col=2)
lines(XX, q2, lwd=2, lty=2, col=2)

deleteGPsep(gpi)
rm(mle)
```

## References

[Page 3 re dummy data function](https://cran.r-project.org/web/packages/tgp/vignettes/tgp2.pdf)

[GPBoost](https://github.com/fabsig/GPBoost)

[Gold price function](https://cran.r-project.org/web/packages/GPfit/GPfit.pdf)
