---
title: "Generalised Additive Models - mgcv (WIP)"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, error = FALSE)
```
<br>

A replication of the ecological models per [Hierarchical generalized additive models in ecology: an introduction with mgcv](https://peerj.com/articles/6876/) using stock data.


<br>

### Packages and data

Load the required packages and the standard data set for analysis

```{r }
library(romerb)
library(mgcv)
library(gratia)

data("stock_data")
fundamental_raw <- stock_data
rm(stock_data)

# Data
df <- fundamental_raw[fundamental_raw$date_stamp == as.Date('2021-06-30'), ]   
df$log_mkt_cap <- log(df$mkt_cap)
df$log_book <- log(-df$total_equity_cln)
df$roe <- df$roe * -1
df <- df[df$date_stamp == as.Date('2021-06-30'), c('symbol','sector','log_book','log_mkt_cap','log_pb','roe','leverage')]
```

<br>

### Hierachical GAM (model type GS)  

Using the factor smooth basis.


```{r }
# Model
gam_mlm1 <- gam(
  log_pb ~ s(roe, k = 5, m = 2) + s(roe, sector, k = 5, m = 2, bs = "fs"),
  data = df, 
  method = "REML"
  )

# Predict
gam_mlm1_pred <- predict(gam_mlm1, se.fit = TRUE)
df$pred <- gam_mlm1_pred$fit
df$sepred <- gam_mlm1_pred$se.fit

# Visualise
ggplot(data = df, aes(x = roe, y = log_pb, group = sector)) +
  facet_wrap(~reorder(sector, as.numeric(sector)), ncol = 4, scales = 'free') +
  geom_ribbon(aes(ymin = pred - 2 * sepred,
                  ymax = pred + 2 * sepred), alpha=0.25) +
  geom_line(aes(y = pred)) +
  geom_point(alpha = 0.3) +
  labs(x = 'Return on equity',
       y = 'Log price / book ratio')
```
<br>

#### Non-hierachical & 3d visualisation 

```{r }
# Model
gam_mlm2 <- gam(
  log_pb ~ s(roe, leverage, bs = "tp", k = 5, m = 2), #+ s(roe, sector, k = 5, m = 2, bs = "fs"),
  data = df, 
  method = "REML"
  )

# Predict
interval <- 30
roe <- with(df, seq(min(roe), max(roe), length = interval))
leverage <- with(df, seq(min(leverage), max(leverage), length = interval))

new_data <- expand.grid(roe = roe, leverage = leverage)

gam_mlm2_pred <- matrix(predict(gam_mlm2, new_data), interval, interval) #, type = "response"


# Visualise 
# https://bikeactuary.com/datasci/gams_scams_pt1
persp(
  x = leverage, 
  y = roe, 
  zlab = 'log_pb',
  gam_mlm2_pred, theta = 225, phi = 20, col = "blue", ticktype = "detailed")
```


<br>

### Single level interaction model  

Using a tensor product smooth.  Modelling market capitalisation on return on equity and book value of equity.

```{r }
# Model
gam_te1 <- gam(
  log_mkt_cap ~ te(roe, log_book, bs=c("cc", "tp"), k=c(10, 10)),
  data = df, method = "REML"
  )

# Visualise
gratia::draw(gam_te1)
```



<br>

Same model, visualise with a 3d plot.  

```{r }
# Predict
interval <- 30
roe <- with(df, seq(min(roe), max(roe), length = interval))
log_book <- with(df, seq(min(log_book), max(log_book), length = interval))

new_data <- expand.grid(roe = roe, log_book = log_book)

gam_te1_pred <- matrix(predict(gam_te1, new_data), interval, interval)

# Visualise
persp(
  x = log_book, 
  y = roe, 
  zlab = 'log_mkt_cap',
  gam_te1_pred, theta = 45, phi = 20, col = "blue", ticktype = "detailed")
```

<br>


Modelling price to book ratio on return on equity and leverage.

```{r, results="hide", message=FALSE, warning=FALSE}
gam_te2 <- gam(
  log_pb ~ te(roe, leverage, bs=c("cc", "tp"), k=c(10, 10)),
  data = df, method = "REML"
)

gratia::draw(gam_te2)
```

<br>

<br>

Same model, visualise with a 3d plot.  

```{r }
# Predict
interval <- 30
roe <- with(df, seq(min(roe), max(roe), length = interval))
leverage <- with(df, seq(min(leverage), max(leverage), length = interval))

new_data <- expand.grid(roe = roe, leverage = leverage)

gam_te2_pred <- matrix(predict(gam_te2, new_data), interval, interval)

# Visualise
persp(
  x = leverage, 
  y = roe, 
  zlab = 'log_pb',
  gam_te2_pred, theta = 225, phi = 20, col = "blue", ticktype = "detailed")
```

<br> 


### Reference  

3d plots with color  

https://stackoverflow.com/questions/22652941/how-to-add-colorbar-with-perspective-plot-in-r
https://stackoverflow.com/questions/31374951/add-points-and-colored-surface-to-locfit-perspective-plot
https://stackoverflow.com/questions/24918604/how-to-have-only-every-other-border-in-a-persp
https://stackoverflow.com/questions/3786189/r-4d-plot-x-y-z-colours
