---
title: "Postgres miscellaneous"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<br>

### Postgres set-up

```{r chunck0}
library(DBI)
library(RPostgres)
library(jsonlite)

# Database connection
config <- jsonlite::read_json('C:/Users/brent/Documents/VS_Code/postgres/postgres/config.json')

con <- DBI::dbConnect(
  RPostgres::Postgres(),
  host      = 'localhost',
  port      = '5432',
  dbname    = 'stock_master_test',
  user      = 'postgres',
  password  = config$pg_password
)
```

<br>

### Insert and update with `insert on conflict`  

Unfortunately Rmarkdown does not allow multiple SQL statements in a single code chuck. Postgres interprets that as submission of multiple commands and throws an error:

`Error: Failed to prepare query: ERROR:  cannot insert multiple commands into a prepared statement`.  

The commented out code is therefore executed in hidden chucks for ease of visualisation. 

<br>  

```{sql chunck1, connection=con}
create temporary table ref_data (
	symbol 			varchar(4) unique
	,company_name 	varchar(50)
	,start_date		date
	,end_date		date
);

--insert into ref_data values ('AAIC', 'American Airlines Group Inc', '2018-06-30', '9998-12-31');
--insert into ref_data values ('AAME', 'Atlantic American Corp', '2020-09-30', '9998-12-31');

--select * from ref_data;
```

```{sql chunck2, connection=con, echo=FALSE}
insert into ref_data values ('AAIC', 'American Airlines Group Inc', '2018-06-30', '9998-12-31');
```

```{sql chunck21, connection=con, echo=FALSE}
insert into ref_data values ('AAME', 'Atlantic American Corp', '2020-09-30', '9998-12-31');
```

```{sql chunck3, connection=con, echo=FALSE}
select * from ref_data;
```

```{sql chunck4, connection=con}
create table new_data (like ref_data);

--insert into new_data values ('BOLT', 'Bolt Biotherapeutics Inc', '2021-03-31', '9998-12-31');
--insert into new_data values ('AAME', 'Atlantic American Corporation', '2020-09-30', '2023-03-03');

--select 'ref_data' as src, r.* from ref_data r
--union all
--select 'new_data' as src, n.* from ref_data n;
```

```{sql chunck5, connection=con, echo=FALSE}
insert into new_data values ('BOLT', 'Bolt Biotherapeutics Inc', '2021-03-31', '9998-12-31');
```

```{sql chunck51, connection=con, echo=FALSE}
insert into new_data values ('AAME', 'Atlantic American Corporation', '2020-09-30', '2023-03-03');
```

```{sql chunck6, connection=con, echo=FALSE}
-- Show data
select 'ref_data' as src, r.* from ref_data r
union all
select 'new_data' as src, n.* from new_data n;
```


```{sql chunck7, connection=con}
insert into ref_data (select * from new_data) 
on conflict (symbol) do update set end_date = excluded.end_date;
```

```{sql chunck8, connection=con}
select * from ref_data
```

As of Postgres 15, the same operation can be executed with `MERGE`

```{sql chunck9, connection=con, eval=FALSE, echo=TRUE}
merge into ref_data r 
using new_data n 
on r.symbol = n.symbol 
when matched then 
	update set end_date = n.end_date
when not matched then 
	insert (symbol, company_name, , start_date, end_date)
	values (n.symbol, n.company_name, n.start_date, n.end_date);
```

```{sql chunck10, connection=con, echo=FALSE}
drop table ref_data
```

```{sql chunck11, connection=con, echo=FALSE}
drop table new_data
```

```{r chunck12, echo=FALSE}
dbDisconnect(con)
```